Simple Mediation Example

### 2017-06-12 mediation.Rmd
simple mediation function now handles covariates.

# simulate strong correlation between x and m

xm_data <- as_tibble(mvrnorm(n, mu = c(6, 4), Sigma = matrix(c(1,0.7,0.7,1), ncol = 2)))
# assign col names
colnames(xm_data) <- c("x", "m")
# check to amek sure correlation is ~70%
cor(xm_data)

# making continous x
sim_data <- xm_data %>% 
  # simulate protective effect between x and y
  mutate(y = rbinom(n, 1, 1/(1+exp(-(1+-0.6*m)))))
```

@@

```{r mediation function}
# define function, mediation (simple function that doesn't have covariate)
mediation <- function(data, y, x, m, y_family, m_family){

 total_effect <- glm(y~x, data, family=y_family)$coefficients[2]
 cde <- glm(y~x+m, data, family=y_family)$coefficients[2]
 nie <- (glm(m~x, data, family=m_family)$coefficients[2]*
         glm(y~x+m, data, family=y_family)$coefficients[3])

@@

```{r testing mediation funciton}
# testing mediation function  
test <- mediation(sim_data,y="y", x="x", m="m", y_family="binomial",
                 m_family="gaussian")

test
```

@@

```{r mediation boot}
# booting 1000 times 
boot_mediation <- sim_data %>% bootstrap(1000) %>% 
  do(tidy(mediation(data=., y="y", x="x", m="m", y_family="binomial",
                 m_family="gaussian")))

>>>>

# simulate strong correlation between x and m 
# also going to add a confounder correlated with both x and m
cov_mat <- matrix(c(1,0.7,0.4,0.7,1,0.3,0.4,0.3,1), ncol=3)
xm_data <- as_tibble(mvrnorm(n, mu = c(6, 4, 10), Sigma = cov_mat))
# assign col names
colnames(xm_data) <- c("x", "m", "c")
# check to amek sure correlation is ~70%
cor(xm_data)

# making continous x
sim_data <- xm_data %>% 
  # simulate protective effect between x and y
  mutate(y = rbinom(n, 1, 1/(1+exp(-(1+-0.6*m + 0.01*c)))))
```

@@

```{r mediation function}
# define function, mediation (simple function that doesn't have covariate)
mediation <- function(data, y, x, m, c, y_family, m_family){

  # if c covariates not specified
  if(missing(c)){
  # define total effect model
  te_mod <- glm(y~x, data, family=y_family)
  # define cde mod
  cde_mod <- glm(y~x+m, data, family=y_family)
  # define x m mod
  m_x_mod <- glm(m~x, data, family=m_family)

  # now if c is specified, adjust for covariates
  } else {  
  te_mod <- glm(as.formula(paste(y, paste(x,c, sep="+"), sep="~")), data,
                family=y_family)

  cde_mod <- glm(as.formula(paste(y, paste(x,m,c, sep="+"), sep="~")), data,
                family=y_family)

  m_x_mod <- glm(as.formula(paste(m, paste(x,c, sep="+"), sep="~")), data,
                family=m_family)
  } # end else

  # fill in estimates from functions
  total_effect <- te_mod$coefficients[2]
  cde <- cde_mod$coefficients[2]
  nie <- m_x_mod$coefficients[2]*cde_mod$coefficients[3]

@@

```{r testing mediation funciton}
# testing mediation function without covariate
test <- mediation(sim_data,y="y", x="x", m="m", y_family="binomial",
                 m_family="gaussian")

test

# test with covariate
test2 <- mediation(sim_data,y="y", x="x", m="m", c="c", y_family="binomial",
                 m_family="gaussian")

test2
```

@@

```{r mediation boot}
# booting 1000 times 
boot_mediation <- sim_data %>% bootstrap(1000) %>% 
  do(tidy(mediation(data=., y="y", x="x", m="m", c="c", y_family="poisson",
                 m_family="gaussian")))
